---
title: "Casestudy1"
author: "Neha Dixit"
date: "6/20/2019"
output: html_document
description: "ABI - Exploratory Data analysis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction
Anheuser-Busch is the leading American brewer holding a 47.6% share of U.S. beer sales to the retailers. The company brews the world’s largest selling beers- Budweiser and budlight and operates many breweries across the U.S.

The U.S. commercial pricing and planning management team at Anheuser-Busch is responsible for tracking the financial performance of sales at retail locations. In order to drive revenues, they are looking to build a model to make recommendations for sales and distributions , such as which brand and how of the product should be sold where.

In order to understand the data and draw insights, SMU consulting has been engaged to perform intial exploratory data analysis on Brewery and beer datasets. 

##

```{r EDA}
########################################################################################################################

## Import essential libraries

library(ggplot2)
library(dplyr)

########################################################################################################################

## Import source breweries and beer data sets

fileLocation <- "https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%207/Breweries.csv"
breweries <- read.csv(url(fileLocation))
head(breweries)

fileLocation <- "https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%207/Beers.csv"
beers <- read.csv(url(fileLocation))
head(beers)

#################################################################################################################
## Initial data analysis

# count rows columns in both dataset
nrow(breweries)
ncol(breweries)
nrow(beers)
ncol(beers)

# Merge beer and brewery dataset
df <- as.data.frame(merge(breweries, beers, by.x = "Brew_ID",by.y= "Brewery_id"))
head(df)

# Look for outliers 
ggplot(df,aes(x=" ", y=df$ABV))+geom_boxplot()+ggtitle("ABV Box plot")+ylab("ABV")
ggplot(df,aes(x=" ", y=df$IBU))+geom_boxplot()+ggtitle("IBU Box plot")+ylab("IBU")

# NA count
na.count <-sapply(df, function(y) sum(length(which(is.na(y)))))
na.count

#################################################################################################################
### Questions

# Q1 How many breweries are present in each state?
# As clearly shown below colorado leads with 47 breweries followed by california at 39 and rest others in deceding order of the brewery count 
breweries %>% count(breweries$State) -> breweries.state

# Rename dataframe column names and display results sorted desc( by Brewery count)
names(breweries.state) <- c("State", "BreweriesCount")
breweries.state[order(-breweries.state$BreweriesCount),]

# Q2 Print the first 6 observations and the last six observations of merged beer and brewery data set 
rbind(head(df),tail(df))

# Q3 Report the number of NA's in each column
# ABV has 62 and IBU has 1005 NA values
na.count <-sapply(df, function(y) sum(length(which(is.na(y)))))
na.count

# Q4 Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare
# Median ABV Content
df %>% select (State, ABV) %>% group_by(State) %>%summarize(median.abv=round(median(ABV, na.rm=TRUE),4)) %>%   arrange(desc(median.abv)) -> abvsummary
data.frame(abvsummary)

# Median IBU Content
df %>% select (State, IBU) %>% group_by(State) %>%summarize(median.ibu=round(median(IBU, na.rm=TRUE),4)) %>%   arrange(desc(median.ibu))  -> ibusummary
data.frame(ibusummary)

# Alcoholic beer by state
ggplot(abvsummary, aes(reorder(State,median.abv), median.abv)) + geom_bar(aes(fill=State),stat="identity") +  coord_flip()+ggtitle("Median ABV by state") +
  xlab(" State \n\n") + ylab("ABV Content")+coord_flip()

# Bitter beers by state
ggplot(ibusummary, aes(reorder(State,median.ibu), median.ibu)) + geom_bar(aes(fill=State),stat="identity") +  coord_flip()+ggtitle("Median IBU by state") +
  xlab(" State \n\n") + ylab("IBU Content")+coord_flip()

# Median ABV Content (top 10 states)
df %>% select (State, ABV) %>% group_by(State) %>%summarize(median.abv=round(median(ABV, na.rm=TRUE),4)) %>%   arrange(desc(median.abv)) %>% slice(1:10)-> abvsummary
data.frame(abvsummary)

# Median IBU Content (top 10 states)
df %>% select (State, IBU) %>% group_by(State) %>%summarize(median.ibu=round(median(IBU, na.rm=TRUE),4)) %>%   arrange(desc(median.ibu))  %>% slice(1:10)-> ibusummary
data.frame(ibusummary)

# Q5 Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
# Colorado
df[which.max(df$ABV),]

# Q6 Summary statistics, histogram and QQplots for AVB and IBU
summary(df$ABV)

ggplot(df,aes(df$ABV))+geom_histogram(fill=("bisque1"),col="blue",breaks=seq(0, .1, by=.01))+xlab("ABV")
ggplot(df,aes(sample=df$ABV))+stat_qq(color="blue")+xlab("ABV")

ggplot(df,aes(df$IBU))+geom_histogram(fill=("bisque1"),col="blue",breaks=seq(0, 150, by=10))+xlab("IBU")
ggplot(df,aes(sample=df$IBU))+stat_qq(color="blue")+xlab("IBU")

# Q7 Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot
# There seems to be some sort of linear relationship ( as ABV increase IBU tends to increase) but since a lot of IBU values are NA it is hard to conclude
plot(df$ABV, df$IBU,pch = 15, col = rainbow(7), main="ABV and IBU Relation", 
  xlab="Alcoholic Beverage content", ylab="Bitterness content",
  xlim=c(0, .1), ylim=c(0, 140))



```
